---
title: 'Suplementary Material 1: Code to Reproduce Analysis and Plots'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!is.null(knitr::opts_knit$get("rmarkdown.pandoc.to"))){
 if (knitr::opts_knit$get("rmarkdown.pandoc.to") == "docx"){
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
} 
}

options(digits = 9)
options(scipen = 9)
options(knitr.kable.NA="")
```

## Introduction

The aim of this document is to display the R code needed the reproduce the findings of the main text. This documents needs the twinR package to be installed to run.

```{r}
#cleanup memory
gcstuff <- gc(verbose=FALSE); rm(gcstuff);
```


```{r}
#get last birth adding function
source("./R/last_birth.R")

#simplified twinR summary tables 
source("./R/twinR_summary.R")

#fix twinR compute predictions to do prediction with no lambda  as well
source("./R/twinR_predictions.R")

#simple convenience functions
source("./R/utils.R")
```


```{r}
## Identify number of CPU cores available for parallel computing,
## note: using a large number may lead RAM to max out, so you may have to adjust
## that according to your infrastructure:
nb_cores <- min(c(50L, parallel::detectCores() - 1))

## Set option in spaMM:
spaMM::spaMM.options(nb_cores = nb_cores)
```


## Data Import Estonia

The Estonian dataset has been formatted to include the same columns as *the data_births_all* dataset from the **twinR** package. The only difference is that the columns *pop* and *monthly* are excluded as these are constant.

```{r}
#Import and preproccess Estonian Data

data_births_monthly_EE <- readRDS("./data/data_births_all_EE.rds")

#the twinR package expects population to be present
data_births_monthly_EE$pop <- "Estonia" 

## Expand the birth level data for the fit of statistical models:
data_births_monthly_EE <- twinR::expand_data(data_births_monthly_EE)

data_births_monthly_EE <- add_last_birth(data_births_monthly_EE)

data_births_monthly_EE_not_last <- data_births_monthly_EE[!data_births_monthly_EE$last,]

#make the aggregates
dmm_EE <- twinR::aggregate_data(data_births_monthly_EE)
dmm_EE$prob_twin <- dmm_EE$twin_total / dmm_EE$births_total

dmm_EE_nl <- twinR::aggregate_data(data_births_monthly_EE_not_last)
dmm_EE_nl$prob_twin <- dmm_EE_nl$twin_total / dmm_EE_nl$births_total
```



## Data Import 9 Western European Populations

```{r}
##Import and pre-proccess twinR package data

## Filter the raw data to only keep data with monthly resolution:
data_births_monthly <- twinR::filter_data(twinR::data_births_all) 

## Expand the birth level data for the fit of statistical models:
data_births_monthly <- twinR::expand_data(data_births_monthly) 

data_births_monthly <- add_last_birth(data_births_monthly)
data_births_monthly_not_last <- data_births_monthly[!data_births_monthly$last,]

dmm_orig <- twinR::aggregate_data(data_births_monthly)
dmm_orig$prob_twin <- dmm_orig$twin_total /dmm_orig$births_total

dmm_orig_nl <- twinR::aggregate_data(data_births_monthly_not_last)
dmm_orig_nl$prob_twin <- dmm_orig_nl$twin_total /dmm_orig_nl$births_total

```


```{r}
#compared to twinR original add data on :
#-age at first birth (AFB)
#-quantiles range mother birth year
#-quantiles range offspring birth year
#-mean + SE total births
all_tbls <- rbind(build_data_summary.table(data_births_monthly),
                  build_data_summary.table(data_births_monthly_not_last),
                  build_data_summary.table(data_births_monthly_EE),
                  build_data_summary.table(data_births_monthly_EE_not_last))
```


```{r}
all_tbls$Dataset  <- rep(c("All births", "Without last birth"),2)
all_tbls$Populations  <- rep(c("9 Western European", "Estonia"),each=2)
all_tbls <- as.data.frame(all_tbls)
row.names(all_tbls) <- paste0(all_tbls$Populations, " populations ",
                              tolower(all_tbls$Dataset)) 
knitr::kable(t(all_tbls),
             caption = paste0("Summary data on the Estonian and other ",
                               "European populations for both ",
                              "all births and wihout last birth for same mothers"))
```



## Fitting models

```{r}
# import the function to do model fit and predictions
source("./R/fit_models.R")
```


### Full Data


```{r}
## Estonia - mother level data
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total"
dmm_EE_fit <- fitPredictions(dmm_EE, formula, predict = T)

knitr::kable(build_fit_summary.table(dmm_EE_fit$fit))
```

```{r}
## Estonia - birth level data
formula <- "twin ~ 1 + poly(cbind(age, parity), 3) + (1|maternal_id)"
dbm_EE_fit <- fitPredictions(data_births_monthly_EE, formula, nb_boot = 0)

knitr::kable(build_fit_summary.table(dbm_EE_fit$fit))
```


```{r}
## TwinR - mother level data
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total + (1|pop)"
dmm_orig_fit <- fitPredictions(dmm_orig, formula, predict=T)

knitr::kable(build_fit_summary.table(dmm_orig_fit$fit))
```

```{r}
## TwinR - birth level data
formula <- "twin ~ 1 + poly(cbind(age, parity), 3) + (1|maternal_id) + (1|pop)"
dbm_orig_fit <- fitPredictions(data_births_monthly, formula, nb_boot = 0)

knitr::kable(build_fit_summary.table(dbm_orig_fit$fit))
```


### No Last Births Data


```{r}
## Estonia
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total"
dmm_EE_nl_fit <- fitPredictions(dmm_EE_nl, formula, predict = T)

knitr::kable(build_fit_summary.table(dmm_EE_nl_fit$fit))
```

```{r}
## Estonia - birth level data
formula <- "twin ~ 1 + poly(cbind(age, parity), 3) + (1|maternal_id)"
dbm_EE_nl_fit <- fitPredictions(data_births_monthly_EE_not_last, formula,
                                predict = T, nb_boot = 0)

knitr::kable(build_fit_summary.table(dbm_EE_nl_fit$fit))
```


```{r}
## TwinR
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total + (1|pop)"
dmm_orig_nl_fit <- fitPredictions(dmm_orig_nl, formula, predict = T)

knitr::kable(build_fit_summary.table(dmm_orig_nl_fit$fit))
```


```{r}
## TwinR - birth level data
formula <- "twin ~ 1 + poly(cbind(age, parity), 3) + (1|maternal_id) + (1|pop)"
dbm_orig_nl_fit <- fitPredictions(data_births_monthly_not_last, formula, nb_boot = 0)

knitr::kable(build_fit_summary.table(dbm_orig_nl_fit$fit))
```

## Plots 


```{r}
library(ggplot2)
#some nice colors
bc <- c("azure4","purple4", "black", "navy", "darkgoldenrod2", "springgreen3", "gray")

#use new base theme that displays also grid lines
source("./R/twinR_theme.R")
```



### Fig 1a: Estonian vs TwinR Full Data

```{r}

fig2_EE_plot_data <- dmm_EE_fit$results
fig2_orig_plot_data <- dmm_orig_fit$results
```

```{r, fig.height=11.7/2, fig.width=8.3}
infoTxt <- "model prediction with 95% CI and data mean SE"
fig2_ext_orig <- ggplot() +  
  geom_line(data=fig2_EE_plot_data,
              aes(y = estimate,  x=births_total, color="Estonia"),  size = 1) +
    stat_summary(data=dmm_EE[dmm_EE$births_total <17, ],
                 aes(x=births_total, y=prob_twin, color="Estonia", fill = "Estonia"),
                 alpha=1,
                 position = position_nudge(x = -0.1),
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_EE_plot_data,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    color="Estonia", fill = "Estonia"),
                alpha = 0.3) + 
    geom_line(data=fig2_orig_plot_data,
             aes(y = estimate,  x=births_total, color="9 W-European"),  size = 1) +
    stat_summary(data=dmm_orig[dmm_orig$births_total <19, ],
                 aes(x=births_total, y=prob_twin,
                     color="9 W-European", fill="9 W-European"),
                 alpha=1,
                 shape = 1,
                 position = position_nudge(x = 0.1),
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_orig_plot_data,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    fill="9 W-European"),
                alpha = 0.1) +
    ggplot2::scale_x_continuous(breaks = 1:18) + 
    ggplot2::scale_y_continuous(breaks = seq(0,0.03, by=0.005)) +
    ggplot2::coord_cartesian() +
    labs(subtitle = paste0("(a) All births, ", infoTxt),
         y="Per-birth twin. prob.",
         x="Maternal total births")
p2 <- fig2_ext_orig + base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") + labs(fill = "population")

p2
```


### Fig 1b: Estonian vs Others No Last Birth


```{r}
fig2_EE_plot_data_nl <- dmm_EE_nl_fit$results
fig2_EE_plot_data_nl$births_total <- fig2_EE_plot_data_nl$births_total + 1
fig2_orig_plot_data_nl <- dmm_orig_nl_fit$results
fig2_orig_plot_data_nl$births_total <- fig2_orig_plot_data_nl$births_total + 1
```


```{r}
dmm_EE_nl_plot <- dmm_EE_nl
dmm_orig_nl_plot <- dmm_orig_nl
dmm_EE_nl_plot$births_total <- dmm_EE_nl_plot$births_total + 1
dmm_orig_nl_plot$births_total <- dmm_orig_nl_plot$births_total + 1
```


```{r, fig.height=11.7/2, fig.width=8.3}
fig2_ext_orig <- ggplot() +  
  geom_line(data=fig2_EE_plot_data_nl,
              aes(y = estimate,  x=births_total, color="Estonia"),  size = 1) +
    stat_summary(data=dmm_EE_nl_plot,
                 aes(x=births_total, y=prob_twin, color="Estonia", fill = "Estonia"),
                 alpha=1,
                 fun.data=mean_se,
                 position = position_nudge(x = -0.1)) +
    geom_ribbon(data=fig2_EE_plot_data_nl,
                aes(y = estimate, x=births_total, ymin = lwr, ymax = upr,
                    color="Estonia", fill = "Estonia"),
                alpha = 0.3) + 
    geom_line(data=fig2_orig_plot_data_nl,
             aes(y = estimate,  x=births_total, color="9 W-European"),  size = 1) +
    stat_summary(data=dmm_orig_nl_plot,
                 aes(x=births_total, y=prob_twin,
                     color="9 W-European", fill="9 W-European"),
                 alpha=1,
                 shape = 1,
                 fun.data=mean_se,
                 position = position_nudge(x = 0.1)) +
    geom_ribbon(data=fig2_orig_plot_data_nl,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    fill="9 W-European"),
                alpha = 0.1) +
    ggplot2::scale_x_continuous(breaks = 1:18, limits = c(1,NA)) + 
    ggplot2::scale_y_continuous(breaks = seq(0,0.03, by=0.005)) +
    ggplot2::coord_cartesian(ylim=c(0,0.03)) +
    labs(subtitle = paste0("(b) Without last birth, ", infoTxt),
         y="Per-birth twin. prob.",
         x="Maternal total births")
p3 <- fig2_ext_orig + base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") + labs(fill = "population")

p3
```



### Figure 2: Model Mean Predictions from Birth Level Data

```{r}
#use new draw_fig_4C function 
source("./R/twinR_theme.R")
```


```{r}
#extract predictions
fig4C_EE_plot_data <- dbm_EE_fit$results
fig4C_EE_plot_data$twin <- fig4C_EE_plot_data$estimates

fig4C_orig_plot_data <- dbm_orig_fit$results
fig4C_orig_plot_data$twin <- fig4C_orig_plot_data$estimates

fig4C_EE_nl_plot_data <- dbm_EE_nl_fit$results
fig4C_EE_nl_plot_data$twin <- fig4C_EE_nl_plot_data$estimates
fig4C_EE_nl_plot_data$parity <- fig4C_EE_nl_plot_data$parity + 1

fig4C_orig_nl_plot_data <- dbm_orig_nl_fit$results
fig4C_orig_nl_plot_data$twin <- fig4C_orig_nl_plot_data$estimates
fig4C_orig_nl_plot_data$parity <- fig4C_orig_nl_plot_data$parity + 1
```

```{r, fig.height=11.7, fig.width=8.3}

rescale_cd <- scale_y_continuous(limits = c(NA, 0.035)) 

fig2b <- draw_fig_4C(fig4C_orig_plot_data) + 
  labs(subtitle = "(b) 9 W-European, all births") 

fig2a <- draw_fig_4C(fig4C_EE_plot_data) + 
  labs(subtitle = "(a) Estonia, all births") 

fig2d <- draw_fig_4C(fig4C_orig_nl_plot_data, y_lims = c(0.01, 0.035)) + 
  labs(subtitle = "(d) 9 W-European, without last b.") 
  
fig2c <- draw_fig_4C(fig4C_EE_nl_plot_data,y_lims = c(0.01, 0.035)) + 
  labs(subtitle = "(c) Estonia, without last birth") 



gridExtra::grid.arrange(fig2a + guides(), fig2b, fig2c, fig2d, ncol=2) 
```


```{r}
#END
```

