---
title: 'Suplementary Material 4: Draw Interbirth Interval Predictons on the Last Birth Omited Data'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!is.null(knitr::opts_knit$get("rmarkdown.pandoc.to"))){
 if (knitr::opts_knit$get("rmarkdown.pandoc.to") == "docx"){
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
} 
}

options(digits = 9)
options(scipen = 9)
options(knitr.kable.NA="")
```

## Introduction

The aim of this document is to display the R code needed the reproduce the original paper figure 4 equivalent on the Estonian dataset where the last births have been omitted.


```{r}
cat("The current directory is: ", getwd())
```

## Get Models

```{r}
fig_data_files <- sapply(LETTERS[1:3], function(l){
  predDir <- "./data/predictions/"
  paste0(predDir, "data_fig_4",l, ".rds")
})
```

As running the *fit_life_histories* function with scenario “**PISH**” will fit all three life history models: *fit_PP*, *fit_IBI*, and *fit_twinning.binary*. The polynomial orders for each of these models will be estimated as the one leading to the best fit between 0 and 6. The *fit_PP* and *fit_IBI* functions will include the variable twin as a predictor. The *fit_twinning.binary* function will include *maternal_id* as a random effect predictor. So, it is equivalent to running each of these three functions with their default settings and the specified scenario. Hence, to produce the figure 4 equivalent we load these fits and get the fitted models from there.

```{r}
if(!all(sapply(fig_data_files, file.exists))){
  fitDir <- "D:/hpc/projects/natCom_twinR_commentary/exports/fitted_models_EE_nl/"
  load(paste0(fitDir, "fits_PISH_obs.rda")) # fit all
  
  data_fig_4A <- twinR::prepare_data_fig_4A(fits$fit_PP)
  gcs <- gc(verbose = F)
  saveRDS(data_fig_4A, paste0(predDir, "data_fig_4A", ".rds"))
  
  data_fig_4B <- twinR::prepare_data_fig_4B(fits$fit_IBI)
  gcs <- gc(verbose = F)
  saveRDS(data_fig_4B, paste0(predDir, "data_fig_4B", ".rds"))
  
  data_fig_4C <- twinR::prepare_data_fig_4C(fits$fit_twinning.binary)
  gcs <- gc(verbose = F)
  saveRDS(data_fig_4C, paste0(predDir, "data_fig_4C", ".rds"))
} else{
  for(l in LETTERS[1:3]){
    assign(paste0("data_fig_4",l), readRDS(fig_data_files[l]))
    }
}

```

## Plot models 

Next we plot the models exactly as in the original manuscript. Note that the actual parity is +1 than the number on the graphs (because the last birth is omitted in this data, so the *parity* axis shows the parity up until the one to last).

```{r}
fig_4A <- twinR::draw_fig_4A(data_fig_4A)
fig_4B <- twinR::draw_fig_4B(data_fig_4B)
fig_4C <- twinR::draw_fig_4C(data_fig_4C)
```

```{r}
twinR::plot_grid(fig_4A, fig_4B, fig_4C,
          labels = "auto", nrow = 1, label_size = 7, align = "v", axis = "lr")
```


```{r}
ggplot2::ggsave(file = "./exports/fig4.pdf", width = 180, height = 55, units = "mm")
```








```{r}
#END
```

